What can Issues on GitHub and Google Code tell us about Bug Hazards in Android and Java?

### 2015-02-15 android-stacks.tex
Finished sectcion IV.A / Root causes
Did extract-command refactoring for the boxed conclusions

Sharpened the first boxed conclusion.

Removed 'evaluation', as this is an empirical study, but not
an evaluation of results.

@@

\emph{\textbf{Identifying the Concerns Related to Root Exceptions.}} To get a broader view of the root exceptions of stack traces,
we performed an empirical evaluation whose goal was to identify the concerns related to the most reported root exceptions.
Besides the exceptions related to programming mistakes mentined before, we also looked for exceptions related to some concerns that are known as sources of faults in mobile development: concurrency ~\cite{ama2012} backward compatibility ~\cite{McDon13}, security ~\cite{enck2011study,was2010} and resource management (IO, Memory, Batery) ~\cite{Zhang12}.

Since it is infeasible to inspect the code responsible for throwning every exception reported in this study,
the concern identification of each exception was based on meaning of the exception type - defined on 
its Javadoc documentaion and on the Java specification. To exemplify: (i) an instance of ArrayOutOfBoundException  
refers to a programming mistake according to its Javadoc; and (ii) the Java specification lists all 
exceptions related to backward compatibility issues~\cite{javaback} (e.g., InstantiationError, VerifyError, IllegalAccessError).

To perform such mapping, we selected a subset of all reported root exceptions. This subset was composed by 100 root exceptions
which were reported on 95\% of all stack traces analyzed in this study. Hence, based on the inspection of the 
Javadoc related to each exception and Java specification, we identified the concern releated to each root exception. 
Table~\ref{tab:tophundrend} illustrates the result of this empirical evaluation. This table also illustrates the exceptions 
that could not be directly mapped to one of the aforementioned concerns, either because they were too general (i.e., java.lang.Exception,
java.lang.RuntimeException, java.lang.Error) or because they were related to other concerns (e.g., specific to an application or given library).

@@

To ensure the quality of the process, three independent coders classified a randomly selected
sample of 25 exception types (from the total 100) using the same list of concerns;
the interater agreement was high (96\%).
We can therefore assume that the categories are descriptive
enough for covering the range of exception types analyzed.

@@

\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
A multitude of programming mistakes were reported as the main causes of exception stack traces.
52\% of all projects reported at least one exception stack trace on which the NullPoiterException
was the root cause.
}
\end{minipage}}

>>>>

\newcommand{\finding}[1]{
\noindent\fbox{%
\begin{minipage}{.97\columnwidth}%
\emph{#1}%
\end{minipage}}
}

@@

\emph{\textbf{Identifying the Concerns Related to Root Exceptions.}} To get a broader view of the root exceptions of stack traces,
we performed a manual inspection in order to identify the underlying
concerns related to the most frequently reported root exceptions.
Besides the exceptions related to programming mistakes mentioned before, we also looked for exceptions related to some concerns that are known as sources of faults in mobile development: concurrency ~\cite{ama2012} backward compatibility ~\cite{McDon13}, security ~\cite{enck2011study,was2010} and resource management (IO, Memory, Batery) ~\cite{Zhang12}.

Since it is infeasible to inspect the code responsible for throwning every exception reported in this study,
the concern identification of each exception was based on intended
meaning of the particular exception type, as defined in 
its Javadoc documentaion and in the Java specification. 
For example: (i) an instance of ArrayOutOfBoundException  
refers to a programming mistake according to its Javadoc; and (ii) the Java specification lists all 
exceptions related to backward compatibility~\cite{javaback}, such as
InstantiationError, VerifyError, and IllegalAccessError.

To perform this concern analysis, we selected a subset of all reported
root exception, consisting of 100 exceptions
reported on 95\% of all stack traces analyzed in this study. Hence, based on the inspection of the 
Javadoc related to each exception and the Java specification, we
identified the underlying concern releated to each root exception. 
Table~\ref{tab:tophundrend} contains the results of this analysis. This table also illustrates the exceptions 
that could not be directly mapped to one of the aforementioned concerns, either because they were too general (i.e., java.lang.Exception,
java.lang.RuntimeException, java.lang.Error) or because they were
related to other concerns (e.g., specific to an application or a given library).

@@

To ensure the quality of the process, three independent coders classified a randomly selected
sample of 25 exception types (from the total 100) using the same list of concerns;
the inter-rater agreement was 96\%.

@@

\finding{%
More than 50\% of the uncaught exceptions are due to errors in
programming logic, with the NullPointerException as most prevalent
exception.
For another 25\% the root cause relates to resource constraints.
}
